---
title: "16S reanalysis"
author: "Hadrien Gourlé"
date: "4/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmarkdown document is part of a reanalysis effort of 13 metabarcoding studies in coral reefs.

The data have been reanalysed with [dada2](https://benjjneb.github.io/dada2/index.html) and this document aims a visualising and interpreting the results.

The code to run the whole pipeline can be found [here](https://github.com/HadrienG/2018_coral_16S)

## Load and tidy the data

```{r load_packages}
library(tidyverse)
```


```{r tidy_function}
#' merge together the metadata, OTU table and taxa table from dada2 into
#' a tidier tibble ready for plotting with ggplot2
tidy_data <- function(seq, tax, metadata) {
    transposed_seq <- as.data.frame(t(seq)) %>%
        as_tibble() %>%
        rownames_to_column(var = 'OTU')
    taxonomy <- as.data.frame(tax) %>%
        as_tibble() %>%
        rownames_to_column(var = 'OTU')
    taxa_table <- inner_join(taxonomy, transposed_seq, by='OTU') %>%
        gather('run', 'count', starts_with('SRR')) %>%
        gather('rank', 'taxonomy', 2:8) %>%
        group_by(taxonomy, run) %>%
        mutate(total = sum(count)) %>%
        ungroup() %>%
        select(-OTU, -count) %>%
        distinct() %>%
        inner_join(metadata, by = 'run') 
    return(taxa_table)
}
```

```{r load_and_tidy}
metadata <- read_csv("data/samples.txt") 
row.names(metadata) <- metadata$run


bioprojects <- metadata %>%
    select(project) %>%
    distinct()
bioprojects
for (bioproject in bioprojects) {
    seqtab <- readRDS(paste0('seqtab_', bioproject, '.Rds'))
    taxa <- readRDS(paste0('taxa_', bioproject, '.Rds'))
    bioproject <- tidy_data(seqtab, taxa, metadata)
}
```

## Tidy the data

TODO:

- [x] add sample info to taxonomy tables
- [ ] merge taxonomy tables

ideally the tidy table should look like

| sample | rank | taxonomy | count |

==> count in seqtab needs to be two rows in taxa

```{r apply_tidy}
map2(data_df$x, data_df$y, tidy_data, metadata = metadata)

for (row in vector) {
    
}
```


## Descriptive statistics

## Is there common ground?
