---
title: "16S reanalysis"
author: "Hadrien Gourlé"
date: "4/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmarkdown document is part of a reanalysis effort of 13 metabarcoding studies in coral reefs.

The data have been reanalysed with [dada2](https://benjjneb.github.io/dada2/index.html) and this document aims a visualising and interpreting the results.

The code to run the whole pipeline can be found [here](https://github.com/HadrienG/2018_coral_16S)

## Load the data and the necessary packages

```{r load_packages}
library(tidyverse)
```

```{r load_data}
r_data_path <- 'results'
r_data_files <- list.files(path = r_data_path, pattern = ".RData",
                           full.names = TRUE)
for (r_data_file in r_data_files) {
    load(r_data_file)
}

metadata <- read_csv("data/samples.txt")
row.names(metadata) <- metadata$run
```

## Tidy the data

TODO:

- [ ] add sample info to taxonomy tables
- [ ] merge taxonomy tables

ideally the tidy table should look like

| sample | rank | taxonomy | count |

==> count in seqtab needs to be two rows in taxa

```{r tidy_data}
# test proj. will apply to all once working
test_seq <- seqtab_PRJNA189184
test_tax <- taxa_PRJNA189184

transposed <- as.data.frame(t(test_seq)) %>%
    as_tibble() %>%
    rownames_to_column(var = 'OTU')
tax <- as.data.frame(test_tax) %>%
    as_tibble() %>%
    rownames_to_column(var = 'OTU')
# View(test_tax)

joined <- tax %>% 
    inner_join(transposed, by='OTU')

View(joined)
```

```{r tidy_data_2}
gathered <- joined %>%
    gather('run', 'count', starts_with('SRR')) %>%
    gather('rank', 'taxonomy', 2:8) %>%
    group_by(taxonomy, run) %>%
    mutate(total = sum(count)) %>%
    ungroup() %>%
    select(-OTU, -count) %>%
    distinct()

View(gathered)
```

```{r tidy_data_3}
final <- gathered %>% inner_join(metadata, by = 'run') 
View(final)
```


## Descriptive statistics

## Is there common ground?
