---
title: "16S reanalysis"
author: "Hadrien Gourlé"
date: "4/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmarkdown document is part of a reanalysis effort of 13 metabarcoding studies in coral reefs.

The data have been reanalysed with [dada2](https://benjjneb.github.io/dada2/index.html) and this document aims a visualising and interpreting the results.

The code to run the whole pipeline can be found [here](https://github.com/HadrienG/2018_coral_16S)

## Load and tidy the data

```{r load_packages}
library(tidyverse)
library(ggsci)
library(RColorBrewer)
library(gridExtra)
library(patchwork)
```

```{r color_palettes}
colors_13 <-  c("#ffa755", "#3e0a7c", "#591300", "#006cbe", "#80002e",
                "#02e0be", "#944400", "#ff70a5", "#00601b", "#d53694",
                "#46c35b", "#b00061", "#ffca7f")
# color <-  grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]
colors_26 <-  c("#ffa755", "#3e0a7c", "#591300", "#006cbe", "#80002e",
                "#02e0be", "#944400", "#ff70a5", "#00601b", "#d53694",
                "#46c35b", "#b00061", "#ffca7f", "#00327f", "#daa61e",
                "#cb729a", "#62ed9b", "#9a0081", "#677300", "#b553c4",
                "#006b3b", "#8770e9", "#c51e38", "#00bcfb", "#cc4128",
                "#fa94ff")
```



```{r tidy_function}
#' merge together the metadata, OTU table and taxa table from dada2 into
#' a tidier tibble ready for plotting with ggplot2
tidy_data <- function(seq, tax, metadata) {
    transposed_seq <- as.data.frame(t(seq)) %>%
        as_tibble() %>%
        rownames_to_column(var = 'OTU')
    taxonomy <- as.data.frame(tax) %>%
        as_tibble() %>%
        rownames_to_column(var = 'OTU')
    taxa_table <- inner_join(taxonomy, transposed_seq, by='OTU') %>%
        gather('run', 'count', starts_with('SRR')) %>%
        gather('rank', 'taxonomy', 2:8) %>%
        group_by(taxonomy, run) %>%
        mutate(total = sum(count)) %>%
        ungroup() %>%
        select(-OTU, -count) %>%
        distinct() %>%
        inner_join(metadata, by = 'run') 
    return(taxa_table)
}
```

```{r load_and_tidy}
metadata <- read_csv("data/samples.txt") 
row.names(metadata) <- metadata$run

full_seqtab <- dir('results', pattern = 'seqtab_', full.names = TRUE) %>%
    set_names() %>%
    map(readRDS)
full_taxa <- dir('results', pattern = 'taxa_', full.names = TRUE) %>%
    set_names() %>%
    map(readRDS)

full_data <- map2(full_seqtab, full_taxa, tidy_data, metadata = metadata)

coral_dataset <- bind_rows(full_data)
```

## Descriptive statistics

### Coral microbiome

- [x] How many NAs?
- [ ] To what is it due? Is there as many in the published papers?

```{r nas, warning = FALSE}
coral_dataset %>% filter(rank == 'Kingdom') %>% {
    ggplot(.) +
        geom_point(aes(project, total, color = taxonomy),
                   position = 'jitter', alpha = 0.8) +
        theme_minimal() +
        scale_color_brewer(palette = 'Set1', na.value = 'grey', name = 'Kingdom') +
        theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
        ggtitle('Distribution of Kingdoms across datasets') +
        xlab('Dataset') +
        ylab('Counts (log scale)') +
        scale_y_log10()
}
ggsave(filename='figures/nas.png', device = 'png', dpi = 600)
```


- [ ] How does the distribution of (i) phyla (ii) families and (iii) genera compare across samples?

```{r}
top_phyla <- coral_dataset %>%
    filter(rank == 'Phylum') %>%
    drop_na() %>%
    group_by(run) %>%
    top_n(5, total) %>%
    filter(total > 20, taxonomy != 'Dinoflagellata', taxonomy != 'Cnidaria',
           taxonomy != 'Arthropoda') %>%
    arrange(total) %>%
    ungroup()

background <- coral_dataset %>%
    filter(rank == 'Phylum') %>%
    drop_na() %>%
    anti_join(top_phyla) %>%
    group_by(run) %>%
    mutate(count = sum(total)) %>%
    select(-total, -taxonomy) %>%
    rename(total = count) %>%
    ungroup() %>%
    add_column(taxonomy = 'Other') %>%
    distinct() %>%
    filter(total > 0)

top_phyla <- top_phyla %>%
    bind_rows(background)

ggplot(top_phyla) +
    geom_boxplot(aes(reorder(taxonomy, total, max), total), alpha = 0.8,
                 color = 'grey30') +
    geom_jitter(aes(reorder(taxonomy, total, max), total, color = project),
                alpha = 0.25, width = 0.2) +
    scale_color_manual(values = colors_13) +
    theme_minimal() +
    coord_flip() +
    ylab('Counts (log scale)') +
    xlab('Most abundant phyla') +
    ggtitle('Distribution of phyla') +
    guides(colour = guide_legend(override.aes = list(alpha = 1))) +
    scale_y_log10()
ggsave(filename='figures/phyla_box.png', device = 'png', dpi = 600)

ggplot(top_phyla) +
    geom_bar(aes(reorder(project, total, max), total, fill = taxonomy),
             stat = 'identity', position = 'fill', alpha = 0.5) +
    scale_fill_manual(values = colors_26) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle('Proportion of phyla per datasets') +
    xlab('Dataset') +
    ylab('Percentage')
ggsave(filename='figures/phyla_bar.png', device = 'png', dpi = 600)
```


- [ ] Is there a common ground? i.e. is there a shared baseline across samples/datasets?
- [ ] Is there a pattern to see with the location and species info?
- [ ] How do the bleached samples look compared to the others?

### Reproducibility

- [ ] How much does the sequencing depth differ across projects?
- [ ] How does the numbers of OTUs found in each study compare with the number of OTUs found in the original papers?
